Lattice Boltzmann study of mass transfer for Bretherton/Taylor bubble train flow

### 2012-05-06 Text/introduction.tex
Style fixes in the first 20% of the paper. Also reformat the whole text so that it does not require wrapping and is easier to read.

\begin{document}
\maketitle
\begin{abstract}
This work presents a thorough procedure of how to determine the volumetric mass transfer
coefficient
in the context of lattice Boltzmann simulations for the Bretherton/Taylor bubble train flow. In this work we address the situation when the hydrodynamic pattern changes from having a vortex in slug ($Ca<0.7$) to not having it ($Ca>0.7$) \cite{giavedoni-numerical}. For the latter case the bubble shape is not symmetric and cannot be approximated through flat surfaces and circumferences as it is done in literature \cite{vanbaten-circular,kreutzer-overview}. For $Ca<0.7$ where there is a vortex in the slug the tracer is well mixed. Thus, it is common to  use periodic boundary conditions  and the inlet/outlet averaged concentration as the characteristic concentration, Eq. \ref{eq:main:definition}.  The later is not valid for flows where tracer is not well mixed, i.e. $Ca>0.7$. This work through examination of different boundary conditions and different characteristic concentration definitions establishes the procedure to determine the volumetric mass transfer coefficient for flows with the capillary number $0.1<Ca<1.0$. With simulations of a few unit cells the connection between time (used in simulations) and space (used in experiments) domains is established. The following boundary conditions were examined: periodic boundary conditions, open boundaries, and a few unit cells open boundary
simulations. It was shown that the time dependent average concentration to be a characteristic
concentration produce robust results, though a bit overpredicted. The inlet/outlet flux averaged
concentration \cite{vanbaten-circular} to be a characteristic concentration produce unreliable results in the range of capillary numbers specified. In comparison with works \cite{vanbaten-circular,kreutzer-overview} where only one unit cell was simulated, a few
unit cells open boundaries conditions resemble physical situation most closely. However, they  are harder to be conducted 
but provide the volumetric mass transfer coefficient based only on the spatial information for the
capillary number $Ca>0.7$. 
%As well, a few unit cells simulations can provide more information about
%mixing in the liquid slug in comparison with the periodic boundary conditions. 
We show
that all presented in literature strategies are extreme limits of one equation,
Eq. \ref{main:main:main}. Finally, simulation results for different Peclet numbers
are compared with analytical predictions presented by \citet{vanbaten-circular} and shows a good
agreement. The work is of use to people performing mass transfer numerical simulations for bubble
train flow in microchannels and/or within the
lattice Boltzmann model context.
\end{abstract}

\section{Introduction}
\label{intro}
The monolith reactors during last years are getting more attention as a promising alternative to slurry
reactors and trickle bed reactors \cite{kreutzer-overview,bercic-mass}. Reactors usually operate in the Bretherton-Taylor regime \cite{bretherton,taylor} which describes a flow through
a liquid medium of equally sized, long air bubbles, see
Fig. \ref{fig:benchmark:hydro}. This flow regime
is characterized by surface tension dominance over inertia and viscous effects, and by
comparatively small gas flow velocities \cite{yue-mass}. Due to surface tension dominance bubble train flow exhibit advantageous properties which can not be achieved
in macroscopic counterparts: liquid thin films \cite{bretherton} between bubbles and walls remarkably enhance mass
transfer from a gas and walls to liquid; the plug flow regime occurring in monolith reactors allows to perform chemical reactions in slugs only \cite{kreutzer-overview}. Moreover, the low slip velocity between gas and liquid is utilized in the experiments to measure liquid velocity \cite{taylor}:  bubbles travelling with approximately the same velocity as liquid can be captured with a camera. Overall, nowadays one can find a large number of applications
of the Bretherton-Taylor bubble train flow: continuous flow analyzers to measure liquid velocity, lung openings, chemical reactions
as hydrogenation of nitroaromatics, 2-ethyl-hexenal, Fischer-Tropsch synthesis, etc. Extensive reviews of \citet{kreutzer-overview,
gupta-review,yue-mass} cover a significant number of applications.

This work is focused on the gas to liquid mass transfer 
study for the Bretherton/Taylor flow. The good understanding of mass transfer and how it depends on the
parameters as the capillary number, the Reynolds number, slug and bubble lengths allows to properly
manufacture a microchannel with necessary properties to ensure that chemical
reactions are performed in the best manner. The mass transfer coefficient characterizing mass transfer is defined as the flux from the surface divided on the concentration difference between the concentration imposed on the body and the characteristic concentration in the domain. The characteristic concentration in the domain significantly depends on underlying hydrodynamics fields.  For example, experimental studies \cite{yue-mass,bercic-mass} show a complex dependency of the mass transfer coefficient on flow parameters: bubble and slug lengths, bubble velocity, which are in turn depend on the capillary number $Ca$ and the Reynolds number $Re$. 
\citet{yue-mass} established an experimental correlation for the volumetric mass transfer coefficient for a bubble train as
a function of the diffusion coefficient, slug and bubble lengths, and bubble velocity: 
\begin{equation}
\vol =\frac{2}{d_h} \Bigl(\frac{D
\ububble}{\lbubble+\lslug}\Bigr)^{0.5}
\Bigl(\frac{\lbubble}{\lbubble+\lslug}\Bigr)^{0.3},
\end{equation}
where $\vol$ is the volumetric mass transfer coefficient, $d_h$ is the hydrodynamic capillary
diameter, $\lbubble$ is the bubble length, $\lslug$ is the slug distance (between bubbles),
$\ububble$ is the bubble velocity, $D$ is the diffusion coefficient. 
%This correlation is within $10\%$ accuracy for experimental
%results for the microchannels with rectangular/square crosssections for the following range of
%parameters: bubble velocity $0.4\,\mathrm{m/s}<\ububble<2\,\mathrm{m/s}$, bubble length
%$1.4<\frac{\lbubble}{d_h}<6.3$, slug length $1<\frac{\lslug}{d_h}<3.2$,
%hydrodynamic diameter $d_h = 400 \mathrm{\mu m}$.

Thus, the understanding of mass transfer for the bubble train flow is not possible without deep understanding of hydrodynamic patterns. There are a number of studies available for the hydrodynamic study of the bubble train flow: experimental \cite{kreutzer-pressure-drop,cerro-space,cerro-bubble-train} and numerical \cite{wang-non-circular,kuzmin-binary3d,giavedoni-numerical,heil-threedim}. For the flow of long bubbles between parallel plates chosen here as the study case it is indicated that  there exists a vortex in the liquid slug for $Ca<0.7$ , see Fig. \ref{fig:streamlines:tweaked:velocity}.  As well, the bubble shape for low capillary numbers ($Ca<0.1$ \cite{cerro-bubble-train}) is symmetric. Thus, in the case of flow between plates for $Ca<0.1\div0.7$ the bubble shape can be represented as two hemicircles and two planar interfaces with the vortex existing in the liquid slug. This is heavily utilized for mass transfer analytical estimations.

As mentioned  the mass transfer coefficient is defined as the mass flux from certain area, Eq. \ref{eq:main:definition}. Thus, analytical estimations \cite{kreutzer-overview,irandoust} are  based on decomposition of the bubble shape. The mass transfer coefficient is calculated through two separate contributions from two films and two hemicircles. For both contributions the Higbie penetration theory \cite{higbie} is utilized, which states that the mass transfer coefficient  for simple geometry flow depends on the average time of liquid packet interaction with a geometrical feature. It can be calculated as $\sqrt{\frac{\pi D}{t_{\mathrm{char}}}}$, where $t_{\mathrm{char}}$ is the interaction characteristic time, $D$ is the diffusion coefficient. As the example for the Higbie penetration theory, the mass  transfer coefficient for the flow of bubbles between parallel plates  is calculated as (similarly to work of \citet{vanbaten-circular}):
\beq
k_L=2 \sqrt{\frac{\pi D}{t_{\mathrm{film}}}}+2 \sqrt{\frac{\pi D}{t_{\mathrm{circle}}}},
\feq
where $t_{\mathrm{film}}=\frac{\lfilm}{\ububble}$ states for the interaction time of liquid traveling through planar part of the bubble, $t_{\mathrm{circle}}=\frac{\pi R_{\mathrm{circle}}}{\ububble}$ is the time which liquid in the slug travels the distance of half bubble circumference.

Despite the simplicity, analytical expressions work well for flows with low capillary numbers $Ca<0.1$ \cite{bercic-mass} where the bubble shape is symmetrical and can be nicely approximated. Moreover, because of the hydrodynamic pattern in the slug, i.e. there is a vortex in slug, one can estimate the time for fluid batch to travel whole circumference.   However, with the change of the capillary number the situation drastically changes. The symmetrical bubble shape changes and resembles a bullet \cite{kuzmin-binary2d}. With the change of the shape and for flows with $Ca>0.7$ there is no vortex in the liquid slug. In this case the Higbie theory fails to estimate contribution from bubble caps. Thus, the need of numerical simulations where all hydrodynamics fields and complicated bubble shapes are taken in the account is obvious.

However, available numerical studies of mass transfer \cite{kreutzer-overview,vanbaten-circular} are lacking the simulation of  bubble shapes. The usual simulation of the mass transfer  is performed as follows: 
\begin{description}
\item[I] The bubble shape is calculated either through analytical correlations \cite{bretherton} or experimental correlations \cite{cerro-bubble-train} without directly resolving bubbles shapes through multiphase simulations.  The expressions for bubble shapes are available only for flows with the capillary number $Ca<0.1$.
\item[II] Hydrodynamics fields are then obtained by performing simulations of one-component flow around the droplet by imposing the bubble velocity on the walls. Thus, the simulations are performed in the reference frame moving with the bubble. The stress free condition is imposed at the bubble surface. 
\item[III]  The mass transfer simulations are  performed in the reference frame moving with the bubble. The saturation concentration is imposed at the bubble surface. Only one unit cell with one bubble in it is used for simulations. Periodic boundary conditions are utilized to determine the volumetric mass transfer coefficient, which is calculated through the following equation \cite{vanbaten-circular}:
\label{main:simulation:equation}
\vol=\frac{\mathrm{\overline{Flux}}}{\cbubble-\langle\coutlet\rangle} \frac{\mathrm{bubble\ surface\ area}}{\mathrm{unit\
cell\ volume}},
where $\langle\coutlet(t)\rangle=\int{C \uoutlet \mathrm{d}A}/\int{\uoutlet\mathrm{d}A}$ is the averaged in space outlet concentration as the function of time used as the characteristic concentration in the definition, Eq. \ref{eq:main:definition}. 
The averaged in time concentration flux ($\mathrm{\overline{Flux}}$) is calculated as the difference between the overall
average concentration in the whole domain ($\langle\coverall\rangle=\int_{V} C \mathrm{d}V /V$)
at time
$t_1$ and at time $t_2$ divided on the time difference $t_2-t_1$. The agreement between numerical
simulations \cite{vanbaten-circular} and correlations of \citet{bercic-mass} was good. 
\end{description}
There is a certain critics towards the presented numerical approaches \cite{vanbaten-circular,kreutzer-overview}. They mainly originate from the bubble shape approximation. It is taken symmetrical, i.e. consisting from hemispheres and cylinder film for the case of flow in circular capillaries. This is valid  for small capillary numbers $Ca<0.1$. As it was discussed, for such capillary numbers in the reference frame moving with the bubble there is a vortex in slug. Thus, the tracer is well mixed in slug.The choice of the characteristic concentration needed for the mass transfer coefficient, Eq. \ref{eq:main:definition}, in this case  is obvious. With small differences in results it can be either averaged in the liquid slug concentration or the outlet space averaged concentration as used in formulation of \citet{vanbaten-circular}. Another critics is towards periodic boundary conditions to calculate the mass transfer coefficient. While it is clear to use periodic boundary conditions for the calculation of hydrodynamics fields for  long bubbles motion in the microchannel, it is not the case for mass transfer simulations. Experimental correlations \cite{bercic-mass} show that the concentration along the streamwise direction changes as the exponential function. However, mass transfer simulations are made only for one unit cell using the periodic boundary conditions with the same concentration at the inlet and at the outlet. The question how  one unit cell simulation corresponds to experimental measurements arises where concentration difference is measured at the distances of at least of a few unit cells difference \cite{bercic-mass}. In other words one needs to understand how the discrete one unit cell simulation corresponds to the continuous picture in experiments where one does not distinguish bubbles but takes measurements of concentration at different locations.

Addressing situations for a rich number of hydrodynamic patterns, shapes, effects for a bubble
train we feel that there is a need to examine more carefully the strategies and assumptions which
stay behind the mass transfer coefficient numerical simulations. We aim at establishing
numerical simulations procedures as to how properly obtain the mass transfer coefficient via a study of different boundary conditions, different characteristic concentration definitions. The case we want to examine is the two-dimensional case of the bubble train flow between parallel plates. The following issues will be addressed:
\begin{description}
\item[I] Applicability of periodic boundary conditions to determine the mass transfer coefficient when there is a change of pattern from having vortex in the slug to not having it. 
\item[II] Validity to use the inlet/outlet averaged or domain averaged concentrations as characteristic concentrations in the definition of the mass transfer coefficient.
\item[III] Transition of the continuous experimental picture to numerical simulations of a few unit cells mass transfer. This is the question about connection between time and space domains. In experiments the characteristic concentration is taken as an average in time, but in numerical simulations \cite{vanbaten-circular} the characteristic concentration is taken as an average in space.
\item[IV] How the change of the hydrodynamic pattern ($Ca>0.7$) influences mass transfer.
\end{description}
Note that the goal of this paper is to establish procedures and boundary conditions used for the mass transfer coefficient determination. Though at the end of the manuscript we present a comparison of the volumetric mass transfer coefficient on the Peclet number with analytical \cite{irandoust} and experimental correlations \cite{yue-mass}, the thorough
determination of the Sherwood number as a function of
other non-dimensional parameters as gas holdup, bubble/slug lenghts, the capillary number is the goal of the future work.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/benchmark_hydro.eps}
\caption{Simplified sketch of the bubble train motion. Using periodic conditions for the velocity
field is natural, but needs validation for the mass transfer. \label{fig:benchmark:hydro}}
\end{figure}

To establish numerical procedures we performed multiphase simulations \cite{kuzmin-binary2d,kuzmin-binary3d} for the range of capillary numbers $Ca=0.1\div 1.0$ to extract bubble shapes. For this range of capillary numbers we are able to capture the bubble shape change and the change of hydrodynamic patterns. The mass transfer simulations were performed with different boundary conditions (open, periodic) and with a few unit cells ($1$ to $10$ unit cells).  The numerical approach we take is the lattice Boltzmann method, a relatively new CFD
competitor emerged during last $20$ years \cite{frisch,mcnamara,HJ,HSB}. During years the
method was applied to simulate not only hydrodynamic problems \cite{yu}, but as well multiphase
flows \cite{Shan-chen:extended,swift,gunstensen}, heat transfer
\cite{yuan-thermal,zhang-thermal}, ferrofluids \cite{dellar-ferro,kuzmin-aniso}.

The mass transfer problems in the lattice Boltzmann framework were mainly addressed in series of
works of Ginzburg and co-authors
\cite{ginzburg-main,ginzburg-boundary-conditions,ginzburg-saturated-flow}. However, all these works
are of general nature to simulate the advection-diffusion equation via the lattice Boltzmann
framework. In comparison, this work focuses on the application side as to establish the procedure of how to obtain the
volumetric mass transfer coefficients for bubble train flow. One should mention the work of
\citet{inamuro-scalar-boundary} about heat and mass transfers in the porous media and the work
of \citet{jos-mass} simulates lateral mixing in cross-channel flow. While two last works are focused
at the particular mass transfer problems, both problems are of homogeneous nature and do not
guide as how to obtain the mass transfer coefficient for non-homogenous problems.

The paper is organized as follows. We start with definitions of the volumetric mass transfer coefficient and apply them 
for the bubble train flow to derive expressions to connect space and time domains. Then the lattice
Boltzmann model used to simulate mass transfer is presented followed by benchmarks. Finally,
numerical simulations of different boundary conditions and a few unit cells simulations for
different hydrodynamic patterns are presented to establish the thorough procedure to determine the
volumetric mass transfer coefficient. The comparison with analytical correlations is also presented.

\section{Mass transfer definitions}
By the definition the mass transfer coefficient from the surface with the imposed constant
concentration $\cbubble$ is the following:
\beq
\label{eq:main:definition}
k_L=\frac{\dot{m}}{P \Delta C},
\feq
where $\dot{m}$ is the mass rate $\Bigl[\frac{kg}{s}\Bigr]$, $P$ is the area of the surface
$\Bigl[m^2\Bigr]$, $\Delta C$ is the concentration difference between the surface and the surrounding medium
$\Bigl[\frac{kg}{m^3}\Bigr]$. Therefore, $k_L$ has a dimension of velocity
$\Bigl[\frac{m}{s}\Bigr]$. Usually, the surrounding medium concentration is taken at the infinite distance
from the bubble. However, in the case of complicated geometries and non-homogeneous concentrations, 
the medium concentration can be the average concentration in the domain or the flux averaged
concentration at the inlet or outlet, etc. Thus, one needs to establish a thorough definition of the volumetric
mass transfer coefficient in the case of complex geometries and non-trivial hydrodynamic velocity patterns.

>>>>

\begin{document}
\maketitle
\begin{abstract}
This work presents a procedure for the determination of the volumetric mass transfer
coefficient in the context of lattice Boltzmann simulations for the Bretherton/Taylor bubble train
flow for capillary numbers $0.1 < Ca < 1.0$. We address the case where the hydrodynamic pattern changes from having a
vortex in slug ($Ca<0.7$) to not having it ($Ca>0.7$) \cite{giavedoni-numerical}.  In the latter case
the bubble shape is asymmetric and cannot be approximated through flat surfaces and circumferences as
is often done in literature \cite{vanbaten-circular,kreutzer-overview}.  When the vortex is present
in the slug, the tracer is well mixed and it is common to use periodic boundary conditions and the inlet/outlet-averaged concentration
% Do we really want equation references in the abstract?
as the characteristic concentration, Eq. \ref{eq:main:definition}.
The latter is not valid for flows where tracer is not well mixed, i.e. $Ca>0.7$.  We therefore
examine various boundary conditions (periodic, open, open with more than 1 unit cell) and definitions of the characteristic concentration.
We show that the time-dependent average concentration produces the most robust results and
that all strategies presented in the literature are extreme limits of one equation,
% Same question applies.
Eq. \ref{main:main:main}. Finally, we show good agreement of simulation results for different Peclet numbers
with analytical predictions \citet{vanbaten-circular}.
\end{abstract}

\section{Introduction}
\label{intro}
Monolith reactors have recently been getting more and more attention as a promising alternative to slurry
reactors and trickle bed reactors \cite{kreutzer-overview,bercic-mass}.  These reactors usually operate in
 the Bretherton-Taylor regime \cite{bretherton,taylor} which describes a flow through
a liquid medium of equally sized, long air bubbles, see
Fig. \ref{fig:benchmark:hydro}. This flow regime
is characterized by surface tension dominance over inertia and viscous effects, and by
comparatively small gas flow velocities \cite{yue-mass}. Due to the dominance of surface tension, the bubble train flow exhibits advantageous
 properties which can not be achieved
in macroscopic counterparts: liquid thin films \cite{bretherton} between bubbles and walls remarkably enhance mass
transfer from gas and walls to liquid; the plug flow regime occurring in monolith reactors allows to perform chemical reactions in slugs only \cite{kreutzer-overview}.
 Moreover, the low slip velocity between gas and liquid is utilized in the experiments to measure liquid velocity \cite{taylor}:  bubbles
 travelling with approximately the same velocity as liquid can be captured with a camera. These properties explain why nowadays one can find a large number of applications
of the Bretherton-Taylor bubble train flow: continuous flow analyzers to measure liquid velocity, lung openings, chemical reactions
such as hydrogenation of nitroaromatics, 2-ethyl-hexenal, Fischer-Tropsch synthesis, etc. The extensive reviews of \citet{kreutzer-overview,
gupta-review,yue-mass} cover a significant number of applications.

This work is focused on the gas to liquid mass transfer 
study for the Bretherton/Taylor flow. A good understanding of mass transfer and how it depends on the
parameters such as the capillary number, the Reynolds number, and slug and bubble lengths allows to properly
manufacture a microchannel with properties necessary to ensure that chemical
reactions are performed in the best possible manner. The mass transfer coefficient is defined as the flux from the 
surface divided by the concentration difference between the concentration imposed on the body and the characteristic concentration in the domain.
 The characteristic concentration in the domain significantly depends on underlying hydrodynamics fields. 
 For example, experimental studies \cite{yue-mass,bercic-mass} show a complex dependency of the mass transfer coefficient on flow parameters:
 bubble and slug lengths, and bubble velocity, which are in turn dependent on the capillary number $Ca$ and the Reynolds number $Re$. 
\citet{yue-mass} established an experimental correlation for the volumetric mass transfer coefficient for a bubble train as
a function of the diffusion coefficient, slug and bubble lengths, and bubble velocity: 
\begin{equation}
\vol =\frac{2}{d_h} \Bigl(\frac{D
\ububble}{\lbubble+\lslug}\Bigr)^{0.5}
\Bigl(\frac{\lbubble}{\lbubble+\lslug}\Bigr)^{0.3},
\end{equation}
where $\vol$ is the volumetric mass transfer coefficient, $d_h$ is the hydrodynamic capillary
diameter, $\lbubble$ is the bubble length, $\lslug$ is the slug distance (between bubbles),
$\ububble$ is the bubble velocity, $D$ is the diffusion coefficient. 
%This correlation is within $10\%$ accuracy for experimental
%results for the microchannels with rectangular/square crosssections for the following range of
%parameters: bubble velocity $0.4\,\mathrm{m/s}<\ububble<2\,\mathrm{m/s}$, bubble length
%$1.4<\frac{\lbubble}{d_h}<6.3$, slug length $1<\frac{\lslug}{d_h}<3.2$,
%hydrodynamic diameter $d_h = 400 \mathrm{\mu m}$.

Thus, the understanding of mass transfer for the bubble train flow is not possible without deep understanding of hydrodynamic patterns.
There are several works studying the hydrodynamic properties of the bubble train flow: 
experimental \cite{kreutzer-pressure-drop,cerro-space,cerro-bubble-train} and numerical \cite{wang-non-circular,kuzmin-binary3d,giavedoni-numerical,heil-threedim}.
 For the flow of long bubbles between parallel plates chosen here as the study case it is indicated that  there exists a vortex 
in the liquid slug for $Ca<0.7$ , see Fig. \ref{fig:streamlines:tweaked:velocity}, and that the bubble shape is symmetric for low capillary 
numbers ($Ca<0.1$ \cite{cerro-bubble-train}).  In the case of flow between plates for $Ca<0.1\div0.7$ the bubble
 shape can be represented as two hemicircles and two planar interfaces with the vortex existing in the liquid slug. This is heavily 
utilized for analytical estimations of mass transfer properties.

Since mass transfer coefficient is defined in terms of a mass flux from a certain area, Eq. \ref{eq:main:definition},
analytical estimations \cite{kreutzer-overview,irandoust} are based on a decomposition of the bubble shape. The mass
 transfer coefficient is calculated through two separate contributions from two films and two hemicircles. For both 
contributions the Higbie penetration theory \cite{higbie} is utilized, which states that the mass transfer coefficient
 for a simple flow geometry depends on the average time of liquid packet interaction with a geometrical feature. It 
can be calculated as $\sqrt{\frac{\pi D}{t_{\mathrm{char}}}}$, where $t_{\mathrm{char}}$ is the characteristic 
time of the interaction, and $D$ is the diffusion coefficient. As an example of the application of the Higbie penetration theory,
 the mass  transfer coefficient 
for the flow of bubbles between parallel plates  is calculated as (similarly to the work of \citet{vanbaten-circular}):
\beq
k_L=2 \sqrt{\frac{\pi D}{t_{\mathrm{film}}}}+2 \sqrt{\frac{\pi D}{t_{\mathrm{circle}}}},
\feq
where $t_{\mathrm{film}}=\frac{\lfilm}{\ububble}$ states for the interaction time of liquid traveling through the planar part
 of the bubble, $t_{\mathrm{circle}}=\frac{\pi R_{\mathrm{circle}}}{\ububble}$ is the time in which liquid in the slug travels the distance of half bubble circumference.

Despite their simplicity, analytical expressions work well for flows with low
capillary numbers $Ca<0.1$ \cite{bercic-mass} where the bubble shape is
symmetrical and can approximated with good precision. Moreover, because of the
hydrodynamic pattern in the slug (i.e. presence of a vortex in slug), one can
estimate the time for a fluid batch to travel whole circumference.   However,
with the increase of the capillary number the situation drastically changes -- the
symmetrical bubble shape is lost and the bubble shape resembles a bullet \cite{kuzmin-binary2d}.
For flows with $Ca>0.7$ there is also no vortex in the liquid slug. In this case the Higbie theory fails
to estimate the contribution from bubble caps, which explains the need to turn to numerical simulations where all
hydrodynamics fields and complicated bubble shapes are taken into account.

Available numerical studies of mass transfer
\cite{kreutzer-overview,vanbaten-circular} lack the simulation of
bubble shapes. The usual simulation of mass transfer is performed as
follows: \begin{description}
 \item[I] The bubble shape is calculated either
through analytical correlations \cite{bretherton} or experimental correlations
\cite{cerro-bubble-train} without directly resolving bubbles shapes through
multiphase simulations.  The expressions for bubble shapes are available only
for flows with the capillary number $Ca<0.1$.
  \item[II] Hydrodynamic fields
are then obtained by performing simulations of one-component flow around the
droplet by imposing the bubble velocity on the walls. Thus, the simulations are
performed in the reference frame moving with the bubble. The stress-free
condition is imposed at the bubble surface.
  \item[III]  The mass transfer
simulations are performed in the reference frame moving with the bubble. The
saturation concentration is imposed at the bubble surface. Only one unit cell
with one bubble in it is used for simulations. Periodic boundary conditions are
utilized to determine the volumetric mass transfer coefficient, which is
calculated through the following equation \cite{vanbaten-circular}:
\begin{equation} \label{main:simulation:equation}
\vol=\frac{\mathrm{\overline{Flux}}}{\cbubble-\langle\coutlet\rangle}
\frac{\mathrm{bubble\ surface\ area}}{\mathrm{unit\ cell\ volume}},
\end{equation}
 where $\langle\coutlet(t)\rangle=\int{C \uoutlet
\mathrm{d}A}/\int{\uoutlet\mathrm{d}A}$ is the space-averaged outlet
concentration as a function of time used as the characteristic concentration
in the definition, Eq. \ref{eq:main:definition}.  The time-averaged
concentration flux ($\mathrm{\overline{Flux}}$) is calculated as the difference
between the overall average concentration in the whole domain
($\langle\coverall\rangle=\int_{V} C \mathrm{d}V /V$) at time $t_1$ and at time
$t_2$ divided by the time difference $t_2-t_1$. The agreement between numerical
simulations \cite{vanbaten-circular} and correlations of \citet{bercic-mass}
was good.  \end{description}

The presented numerical approaches \cite{vanbaten-circular,kreutzer-overview}.
can be critized on a number of points. They mainly
originate from the bubble shape approximation, which is taken to be symmetrical, i.e.
consisting of hemispheres and cylinder film for the case of flow in circular
capillaries. This is valid  for small capillary numbers $Ca<0.1$. As previously
discussed, for such capillary numbers the tracer is well mixed in the slug and 
the choice of the characteristic concentration needed for the mass transfer
coefficient, Eq. \ref{eq:main:definition} is obvious. With small
differences in results it can be either averaged in the liquid slug
concentration or the outlet space-averaged concentration as used in the formulation
of \citet{vanbaten-circular}.

While it clear that periodic boundary conditions can be employed for the calculation
 of hydrodynamic fields, the same does not apply to the mass transfer coefficient.
simulations. Experimental correlations \cite{bercic-mass} show that the
concentration along the streamwise direction changes exponentially with distance.
Mass transfer simulations however, are made only for one unit cell
using the periodic boundary conditions with the same concentration at the inlet
and at the outlet. The question how one unit cell simulation corresponds to
experimental measurements arises where concentration difference is measured at
the distances of at least of a few unit cells \cite{bercic-mass}. In
other words one needs to understand how the discrete one unit cell simulation
corresponds to the continuous picture in experiments where one does not
distinguish bubbles but takes measurements of concentration at different
locations.

Addressing situations for a rich number of hydrodynamic patterns, shapes, and
effects for a bubble train, we feel that there is a need to examine more
carefully the strategies and assumptions behind the numerical calculations
of the mass transfer coefficient. We aim at establishing clear
procedures as to how properly obtain the mass transfer coefficient via a study
of different boundary conditions and different definitions of characteristic concentration
The case we want to examine is a two-dimensionalbubble train flow between parallel plates.
We address the following issues:
 \begin{description}
 \item[I] Applicability of periodic boundary
conditions to determine the mass transfer coefficient when the vortex in the slug
disappears.
 \item[II] Validity of the inlet/outlet-averaged or domain-averaged concentrations as
characteristic concentrations in the definition of the mass transfer
coefficient.
 \item[III] Transition of the continuous experimental picture to
numerical simulations of a few unit cells, the issue of correspondence
between space averages (simulations \cite{vanbaten-circular}) and time averages (experiment).
 \item[IV] Influence of the hydrodynamic pattern in flows with $Ca>0.7$ on the mass transfer.
\end{description}
In addition, at the end of the manuscript we present a comparison of dependence the volumetric
mass transfer coefficient on the Peclet number with analytical \cite{irandoust} and experimental correlations
\cite{yue-mass}.  The thorough determination of the Sherwood number as a
function of other non-dimensional parameters such as gas holdup, bubble/slug
lenghts, the capillary number is the goal of future work.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/benchmark_hydro.eps}
\caption{Simplified sketch of the bubble train motion. Using periodic
conditions for the velocity field is natural, but needs validation for the mass
transfer.
\label{fig:benchmark:hydro}}
\end{figure}

To establish numerical procedures we performed multiphase simulations
\cite{kuzmin-binary2d,kuzmin-binary3d} for the range of capillary numbers
$Ca=0.1\div 1.0$ to extract bubble shapes. For this range of capillary numbers
we are able to capture the bubble shape change and the change of hydrodynamic
patterns. The mass transfer simulations were performed with different boundary
conditions (open, periodic) and with a few unit cells ($1$ to $10$ unit cells).
As our numerical approach we take the lattice Boltzmann method, a relatively
new CFD competitor developed during last $20$ years
\cite{frisch,mcnamara,HJ,HSB}. This method was successfully applied to simulate
not only simple hydrodynamic problems \cite{yu}, but also multiphase
flows \cite{Shan-chen:extended,swift,gunstensen}, heat transfer
\cite{yuan-thermal,zhang-thermal}, and ferrofluids \cite{dellar-ferro,kuzmin-aniso}.

The mass transfer problems in the lattice Boltzmann framework were mainly
addressed in series of works of Ginzburg and co-authors
\cite{ginzburg-main,ginzburg-boundary-conditions,ginzburg-saturated-flow}.
In contrast to these works whose focus was on simulating the
advection-diffusion equation via the lattice Boltzmann framework, we concentrate
on the application side.  One should also mention the work of \citet{inamuro-scalar-boundary}
about heat and mass transfers in porous media and the work of
\citet{jos-mass} simulating lateral mixing in cross-channel flow.  The last two
works are focused on problems of homogeneous nature and do not provide
guidance as how to obtain the mass transfer coefficient for non-homogenous cases.

The paper is organized as follows. We start with definitions of the volumetric mass transfer coefficient and apply them 
for the bubble train flow to derive expressions to connect the space- and time-averages.  Then, the lattice
Boltzmann model used to simulate mass transfer is presented, and followed by benchmarks. Finally,
numerical simulations of different boundary conditions and simulations spanning a few unit cells for
different hydrodynamic patterns are presented to establish the procedure to determine the
volumetric mass transfer coefficient. The comparison with analytical correlations is also presented.

\section{Mass transfer definitions}
By definition, the mass transfer coefficient from the surface with the imposed constant
concentration $\cbubble$ is the following:
\beq
\label{eq:main:definition}
k_L=\frac{\dot{m}}{P \Delta C},
\feq
where $\dot{m}$ is the mass rate $\Bigl[\frac{kg}{s}\Bigr]$, $P$ is the area of the surface
$\Bigl[m^2\Bigr]$, and $\Delta C$ is the concentration difference between the surface and the surrounding medium
$\Bigl[\frac{kg}{m^3}\Bigr]$. Therefore, $k_L$ has a dimension of velocity
$\Bigl[\frac{m}{s}\Bigr]$. Usually, the surrounding medium concentration is taken at the infinite distance
from the bubble. However, in the case of complicated geometries and non-homogeneous concentrations, 
the medium concentration can be the average concentration in the domain or the flux averaged
concentration at the inlet or outlet, etc. Thus, one needs to establish a thorough definition of the volumetric
mass transfer coefficient in the case of complex geometries and non-trivial hydrodynamic velocity patterns.

There exist a number of methods to estimate the mass transfer coefficient $k_L$. We first examine the
theoretical definitions of the mass transfer in the case of point mass sources.
\subsection{Point mass sources}
In what follows we will present three approaches to calculate point  mass transfer
coefficients (by point source we assume the source with an infinitesimally small surface area $P$):
\begin{enumerate}
\item
Let us look at the infinitesimal small domain of volume $A \Delta x$, not
moving and with a point mass source.  The concentration difference can be then found as $\Delta C =
\cstar -
C(t)$, where $\cstar$ is the imposed point source concentration, and $C(t)$ is
the time-dependent concentration, which does not depend on the location due to the assumption of
homogeneity. One can therefore write a time-dependent ordinary differential equation for the
concentration in the domain:
\beq
\dot{m}= A \Delta x \frac{\mathrm{d}C}{\mathrm{d} t} = k_L P (\cstar-C(t)), 
\feq
with the initial condition $C(0)=0$
The solution can be found as:
\beq
C(t)= \cstar (1-\exp(-\vol t )), 
\feq
where $\vol$ is the volumetric mass transfer coefficient defined as:
\beq
\vol=k_L \frac{P}{A \Delta x}=k_L \frac{P}{V},
\feq
where $P$ is the source surface, $V$ is the unit cell volume.
\item
Let us predict mass transfer in a liquid moving with the velocity
$U$, see Fig. \ref{fig:moving:frame}. 
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/mass_transfer.eps}
\caption{The mass transfer in a moving liquid. \label{fig:moving:frame}}
\end{figure}

If one can assume that the point mass sources are distributed in the whole
medium, the mass accumulated in the unit volume $V=A \Delta x$ can be
calculated as the difference of mass fluxes entering and leaving domain $U
\bigl(C(x+\Delta x)-C(x)\bigr)$. The accumulated mass should be proportional to
the mass transfer coefficient:
\beq
U \bigl(C(x+\Delta x)-C(x)\bigr)=k_L P \bigl(\cstar-C(x)\bigr), 
\feq 
giving the same equation but only in the coordinate domain:
\beq
%\label{concentration:domain:coordinate}
C(x)= \cstar \Bigl(1-\exp\bigl(-k_L a \frac{x}{U} \bigr)\Bigr).
\label{main:mass:transfer:expression} 
\feq
Note that the concentration $C(x)$ does not depend on time.

\item If one transfers to the frame moving with the liquid velocity $U$, the situation will be
the same as in the first case. However, one can connect the time and spatial location with the
velocity $U$ ($t=\frac{x}{U}$) to obtain the same equation as in the case 2.
\end{enumerate}

\subsection{Bubble train}
In the application to the bubble train flow it is useful to think of one bubble to be a point
source to make all considerations above valid. For example, the expression
(\ref{main:mass:transfer:expression}) was used in the experiments by
\citet{bercic-mass}. However, one should be accurate with the definition of velocities because two
different phases co-exist in the bubble train flow. Usually, one can take the velocity $U$ to be
a bulk velocity or $U=\ugas+\uliq$, where $\ugas$ and $\uliq$ are liquid and gas
superficial velocities, correspondingly.
